The Name of the Title is Hope

### 2024-10-04 submission_JNE/source/input/5_user_study_and_methods.tex
page and jne submission

In the user study, we compared three conditions of agency experience. With the first two conditions, INTENTION and EXTERNAL we queried users at the two edges of agency experience. In EXTERNAL, participants had no intention to move and no control over their movement. On the other hand, in INTENTION, participants were acting as they would in their day-to-day lives, fully in control. We then investigated in a third condition where our AUGMENTED prototype sits on the spectrum between INTENTION and EXTERNAL, in other words, whether it preserved agency and to what degree. To answer our question, we employed a mixed methods approach including a psychometric test of intentional binding, one standardized question, and a qualitative (phenomenological) exit interview.

To assist readers in replicating our prototype and experiment, we provide the necessary technical details, the complete source code for the BCI and the experiment, the collected data, as well as all the analysis scripts\footnote{[anonymized]}

\subsection{Participants}
Eleven participants (M = 29.9 years, SD = 4) were recruited from our local institution and through the institute's online participant pool. Participants were compensated with course credit or 12 Euro per hour of study participation. Prior to their participation, they were informed of the nature of the experiment, recording, and anonymization procedures and signed a consent form. The experiment was approved by the ethics committee of [anonymized]. One participant had to be excluded from further data analyses due to significant deviations from the instructions in the execution of the task. Precisely, they did not initiate their movements at their own volition but rather immediately at the onset of each trial, thereby violating the task instruction of waiting 2-3 s after trial onset, see task description below for more detail.
% All participants were right-handed

@@

\indent\textbf{(4) Experiment Presentation and Collection of Behavioral Responses.} An Acer Group (Acer Inc, Taipeh, Taiwan) tablet was used to present the task to participants and record their behavioral responses. In addition to the tablet, we used an external keyboard to allow users to input their timing judgments.

\subsection{Experimental Task, Design and Procedure}
Participants performed a simple tapping task in three conditions with 75 trials each: INTENTION, EXTERNAL, and AUGMENTED

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/method.png}
    \caption{Diagram summarizing: (a) signal processing and event labeling from muscle activity, (b) extraction of linear slope features from labeled brain signals and (c) real-time application of the BCI gating the muscle stimulation hardware.}
    \label{fig:method}
\end{figure*}

@@

During INTENTION, participants were equipped with EMG sensors instead of EMS electrodes. Participants' reaction times in INTENTION were used to select stereotypical reaction times for the EXTERNAL condition. At the end of INTENTION, EMG electrodes were exchanged for EMS electrodes at the identical location on the forearm.

\indent\textbf{EXTERNAL.} The task structure was identical to INTENTION, however, participants were now instructed to hold and wait for the muscle stimulation to move their finger thereby eliciting the screen tap. The timing for the EMS trigger was taken by randomly choosing a time between the 5th and 95th percentile of their actual individual reaction time in the INTENTION condition.

@@

The order of the conditions was not pseudo-randomized since training data obtained in INTENTION was required for both EXTERNAL and AUGMENTED. Furthermore, AUGMENTED was always the last condition, thereby allowing for a prolonged interview immediately after the experience with the prototype.

\subsection{Brain-Computer Interface}\label{BCI}
The data obtained in INTENTION was used to train the BCI. We utilized the [anonymized] and the EEGLAB~\cite{Delorme2004-sn} toolbox inside the MATLAB (The MathWorks Inc. Natick, MA, USA) environment for preprocessing both EEG and EMG data. First, to generate behavioral labels at a high temporal resolution for the EEG-based classifier, we leveraged EMG data from the flexor digitorum profundus. EMG amplitudes were band-pass filtered from 20 to 100 Hz and subsequently squared. Next, to label the time of movement onset, the EMG data was averaged across trials for the second preceding the screen tap. From this averaged data, the first sample where the EMG amplitude exceeded the 95th percentile was selected as the time of movement onset, see~\ref{fig:method} a.
% TODO remove anonymization bemobil pipeline

% TODO/Ignore Next, noisy data segments were determined. 
% A segment was rejected as an outlier using Matlab's \textit{isoutlier} function using default parameters. % from isoutlier documentation (mean segment amplitude exceeding 3 scaled median absolute deviation)

@@

\subsubsection{Training of EEG classifier}
To classify EEG data, a linear discriminant analysis (LDA) with shrinkage regularization (automatic shrinkage using the Ledoit-Wolf lemma~\cite{Ledoit2004-bi}) was trained for each participant individually. As single-trial features, the (linear) slope coefficient was obtained for both \textit{idle} and \textit{pre-movement} segments by fitting a linear regression. The classifier was then trained using the slope coefficient of the selection of the most discriminative channels as features, generating a feature vector in the dimension of channels that were kept for classification, see~\ref{fig:method} b. We purposefully constrained the dimensionality of the feature vector to avoid over-fitting and decrease the computational load.

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/eeg_results_new.pdf}
    \caption{(a) Grand average event-related potential (ERP) at electrode Cz of pre-movement EEG data epochs preceding the last second before a movement, idle epochs of the same trial are plotted alongside but originate from a different time window, see section~\ref{eeg_methods}; (b) Bottom: Slope features for both idle and pre-movement classes at electrode Cz. Top: Scalp maps of slope values for both classes and all channels. (c): Mean and per participant ROC curves. Dotted line at 15 \% false positive rate indicate the selected threshold for the real-time application.}
    \label{fig:EEG_results}
\end{figure*}

\subsubsection{Real-time application and EMS control}
During real-time application, the EEG data was buffered for the last second for the selected discriminative channels. The data was band-pass filtered analogously to the training data from 0.1 to 15 Hz. Next, the slope was computed. This procedure ran at an update rate of 10 Hz, hence every 100 ms a new prediction was obtained from the classifier. To smooth the prediction output to reduce false predictions due to unlikely peaks, the predicted probability for the \textit{pre-movement} class was smoothed by averaging current and the preceding prediction with a weighting (weight of .3 for the preceding, and .5 for the current prediction). The weights were obtained through trial-and-error during piloting. Then, with a 10 Hz update rate, this smoothed probability, as well as the predicted class for the current frame were gating the EMS switch: when the probability exceeded the threshold and the currently predicted class was \textit{pre-movement}, the switch was opened for 0.5 s, see~\ref{fig:method} c.

@@

\subsection{Measures of Agency Experience: Intentional Binding, Question \& Interview}
To assess if SoA was preserved, an intentional binding measure, a standardized question, and an interview following the AUGMENTED block were investigated. After each condition we prompted participants to rate their experienced agency on a 7-point Likert scale with the statement ``It felt like I was in control of the movements during the task.'', the item was copied from~\cite{Bergstrom2022-fb}. Following AUGMENTED we interviewed users about their experience working with the system. After prompting users to recall their experience and summarize what their task had been, we set the focus to the tapping movement and asked them to ignore the time estimation task in the questions that followed. We entered the open part of the interview by asking: ``What did the system do?'' followed up by ``What was the difference between the three conditions''. After some time, and depending on their answers, we reset the focus to the AUGMENTED condition and asked ``How often was the system active?'' followed by ``What do you think caused the actions of the system?''. This was then followed up by an `open' interview in which we frequently asked `how' and `why' questions to inquire about the user's experience.

% add back in after anonymization: All interviews were manually transcribed and then automatically translated using }
We analyzed the interviews by loosely following~\citet{Mayring2015-pp}. All interviews were manually transcribed and translated to English using DeepL\footnote{https://www.deepl.com/translator} (DeepL SE, Cologne, Germany). Before screening the texts, two experts clustered the responses into 3 clusters: First, ``Functionality'' referred to what participants attributed the source of the stimulation. Next, we clustered responses according to ``Guessed percentage'' of correct interaction, i.e., participants' estimate of how well the system was aligned with their intention. For the cluster ``Correct Interaction'', we specifically queried participants to recall the moments where the stimulation felt in line with their intention to move, then we clustered their responses into sentiments with positive and negative valence.

\subsection{Statistical Analyses}
To confirm and demonstrate the discriminative power of the EEG features, we plotted the amplitude time course of electrode Cz between \textit{pre-move} and \textit{idle} epochs. Electrode Cz was chosen for exemplary presentation, as it is frequently reported in studies on the RP, since it captures neural activity originating from the sensorimotor cortex. Next, the slope coefficients were extracted and a paired t-test was conducted. Ultimately, to assess the classifier performance, we calculated the F1 score, i.e. the harmonic mean of precision and recall, and plotted the ROC, see figure~\ref{fig:EEG_results} a to c.

@@

In line with the literature on intentional binding, we hypothesized that the time intervals should be underestimated for the INTENTION and the AUGMENTED conditions. This should not be the case for the EXTERNAL condition where users had no intention to move. Hence, when binding occurs, the intervals should be underestimated, hinting at a higher SoA.

To test this, we fitted a linear mixed effects model with \textit{condition} (INTENTION, EXTERNAL, AUGMENTED) as a fixed effect and `participandID' as a random effect. The model was specified as `time interval $\sim$ condition + (1|participantID)' and fit using the `lme4' package~\cite{Bates2015-bh}. We included all single trial data points in the analyses and did not aggregate before fitting the model as fitting the random effect `participantID' corrects for the correlation of the repeated measurements within participants. We obtained a test statistic by calculating likelihood-ratio tests comparing the full model as specified above against the null model `time interval $\sim$ 1 + (1|participandID)'. All parameters were estimated by maximum likelihood estimation~\cite{Pinheiro2006-bk}. We computed post-hoc pair-wise tests for `condition' corrected for multiple comparisons (Tukey method) using the emmeans package~\cite{Lenth2020-xk}.

Next, we hypothesized that subjective ratings of control over the tap movement are comparable between INTENTION and AUGMENTED and lower in the EXTERNAL condition. Again, we fitted a linear mixed effects model with \textit{condition} as a fixed effect and participant ID as a random effect. Coefficients were assessed in the same way as for the intentional binding parameter above.

In short, we tested two main hypotheses with regard to the agency experience in our three experimental conditions: (1) participants underestimate the tone delay when acting intentionally. Adding EMS in line with participants' intention to move does not affect this underestimation. (2) The subjective feeling of control is comparable between INTENTION and AUGMENTED conditions. EXTERNAL should decrease the feeling of control significantly. Additionally, we report the clustered interviews anecdotally.

>>>>

We compared three conditions of agency experience. With the first two conditions, INTENTION and INVOLUNTARY we queried users at the two edges of agency experience. In INVOLUNTARY, participants had no intention to move and no control over their movement, much like in a bystander scenario. On the other hand, in INTENTION, participants were acting as they would in their day-to-day lives, fully in control and with volition. We then investigated in a third condition where our AUGMENTED prototype sits on the spectrum between INTENTION and INVOLUNTARY, in other words, whether it preserved agency and to what degree. To answer our question, we employed a mix of qualitative and quantitative methods, including a psychometric test of intentional binding, one standardized question, a qualitative (phenomenological) exit interview, and an exploratory EEG analyses of prediction errors following muscle stimulation.

% To assist readers in replicating our prototype and experiment, we provide the necessary technical details, the complete source code for the BCI and the experiment, the collected data, as well as all the analysis scripts\footnote{[anonymized]}
% https://github.com/lukasgehrke/2021-fastReach

\subsection{Participants}
Eleven participants (M = 29.9 years, SD = 4) were recruited from our local institution and through the institute's online participant pool. Participants were compensated with course credit or 12 Euro per hour of study participation. Prior to their participation, they were informed of the nature of the experiment, recording, and anonymization procedures and signed a consent form. The experiment was approved by the local ethics committee of the Department of Psychology and Ergonomics at the TU Berlin (Ethics protocol approval code: BPN\_GEH\_2\_230130). One participant had to be excluded from further data analyses due to significant deviations from the instructions in the execution of the task. Precisely, they did not initiate their movements at their own volition but rather immediately at the onset of each trial, thereby violating the task instruction of waiting 2-3 s after trial onset, see task description below for more detail.

@@

\indent\textbf{(4) Experiment Presentation and Collection of Behavioral Responses.} An Acer Group (Acer Inc, Taipeh, Taiwan) tablet was used to present the task to participants and record their behavioral responses. In addition to the tablet, we used an INVOLUNTARY keyboard to allow users to input their timing judgments.

\subsection{Experimental Task, Design and Procedure}
Participants performed a simple tapping task in three conditions with 75 trials each: INTENTION, INVOLUNTARY, and AUGMENTED

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/method.png}
    \caption{Diagram summarizing(a) signal processing and event labeling from muscle activity, (b) extraction of linear slope features from labeled brain signals, and (c) real-time application of the BCI gating the muscle stimulation hardware.}
    \label{fig:method}
\end{figure*}

@@

During INTENTION, participants were equipped with EMG sensors instead of EMS electrodes. Participants' reaction times in INTENTION were used to select stereotypical reaction times for the INVOLUNTARY condition. At the end of INTENTION, EMG electrodes were exchanged for EMS electrodes at the identical location on the forearm.

\indent\textbf{INVOLUNTARY.} The task structure was identical to INTENTION, however, participants were now instructed to hold and wait for the muscle stimulation to move their finger thereby eliciting the screen tap. The timing for the EMS trigger was taken by randomly choosing a time between the 5th and 95th percentile of their actual individual reaction time in the INTENTION condition.

@@

The order of the conditions was not pseudo-randomized since training data obtained in INTENTION was required for both INVOLUNTARY and AUGMENTED. Furthermore, AUGMENTED was always the last condition, thereby allowing for a prolonged interview immediately after the experience with the prototype.

\subsection{Brain-Computer Interface}\label{BCI}
The data obtained in INTENTION was used to train the BCI. For processing EEG and EMG data from INTENTION, we utilized the EEGLAB~\cite{Delorme2004-sn} toolbox with wrapper functions from BeMoBIL-pipeline~\cite{Klug2022-lc} running in the MATLAB (The MathWorks Inc. Natick, MA, USA) environment. First, to generate behavioral labels at a high temporal resolution for the EEG-based classifier, we leveraged EMG data from the flexor digitorum profundus. EMG amplitudes were band-pass filtered from 20 to 100 Hz and subsequently squared. Next, to label the time of movement onset, the EMG data was averaged across trials for the second preceding the screen tap. From this averaged data, the first sample where the EMG amplitude exceeded the 95th percentile was selected as the time of movement onset, see~\ref{fig:method}a.

% Ignore Next, noisy data segments were determined. 
% A segment was rejected as an outlier using Matlab's \textit{isoutlier} function using default parameters. % from isoutlier documentation (mean segment amplitude exceeding 3 scaled median absolute deviation)

@@

\subsubsection{Training of EEG classifier}
To classify EEG data, a linear discriminant analysis (LDA) with shrinkage regularization (automatic shrinkage using the Ledoit-Wolf lemma~\cite{Ledoit2004-bi}) was trained for each participant individually. As single-trial features, the (linear) slope coefficient was obtained for both \textit{idle} and \textit{pre-movement} segments by fitting a linear regression. The classifier was then trained using the slope coefficient of the selection of the most discriminative channels as features, generating a feature vector in the dimension of channels that were kept for classification, see~\ref{fig:method}b. We purposefully constrained the dimensionality of the feature vector to avoid over-fitting and decrease the computational load.

@@

\subsubsection{Real-time application and EMS control}
During real-time application, the EEG data was buffered for the last second for the selected discriminative channels. The data was band-pass filtered analogously to the training data from 0.1 to 15 Hz. Next, the slope was computed. This procedure ran at an update rate of 10 Hz, hence every 100 ms a new prediction was obtained from the classifier. To smooth the prediction output to reduce false predictions due to unlikely peaks, the predicted probability for the \textit{pre-movement} class was smoothed by averaging the current and the preceding prediction with a weighting (weight of .3 for the preceding, and .5 for the current prediction). The weights were obtained through trial-and-error during piloting. Then, with a 10 Hz update rate, this smoothed probability, as well as the predicted class for the current frame were gating the EMS switch: when the probability exceeded the threshold and the currently predicted class was \textit{pre-movement}, the switch was opened for 0.5 s, see figure~\ref{fig:method}c.

@@

\subsubsection{Post-hoc evaluation of real-time performance}
To not break the user's focus on the task at hand, (subjective) labels were not obtained during the real-time application in AUGMENTED. In other words, participants were not directly queried, e.g. via questionnaires, to judge whether the stimulation in the current trial was in line with their intention or not. Without these labels, a post-hoc analyses of the binary classifiers performance was not possible, i.e. to ascertain for example false positive stimulations. However, we explored an alternate approach to estimate (subjective) labels that does not break the user's task immersion. We investigated prediction errors in response to the movement onset through event-related potentials (ERPs) at fronto-central electrode FCz. Previously, prediction error ERPs at, among others, electrode FCz have been shown to be one suitable candidate feature to detected breaks in (task) immersion, such as when perceiving glitches in VR~\cite{Gehrke2019-og, Gehrke2022-kz, Gehrke2024-xq, Terfurth2024-kh, Si-mohammed2020-ru}. In BCIs, these ERPs are frequently leveraged to correct system errors~\cite{Zander2016-ed}. In their work,~\citet{Zander2016-ed} demonstrated a BCI to decode a user’s intended cursor movement direction on a 6 × 6 grid. The system regularly probed the user by observing the EEG response to random cursor movements. How severely the random dot movement violated the user’s intention was directly reflected in the ERP activity. We hypothesized to find similar ERP signatures when the stimulation misaligned with the user's prediction.

Since this analyses was conducted post-hoc, it allowed for more signal processing in comparison to the feature extraction from INTENTION when participants were waiting to start the next block. With the goal to best recover the prediction error ERP, we again applied `BeMoBIL-pipeline'~\cite{Klug2022-lc} wrapper functions of EEGLAB~\cite{Delorme2004-sn}: After removing non-experiment segments at the beginning and end of the concatenated recording from all three conditions, EEG data was re-sampled to $250~Hz$. Next, bad channels were detected using the `FindNoisyChannel' function, which selects bad channels by amplitude, the signal-to-noise ratio, and correlation with other channels~\cite{Bigdely-Shamlo2015-ds}. Rejected channels were then interpolated while ignoring the EOG channel, and finally re-referenced to the average of all channels, including the original reference channel FCz. After applying a high-pass filter at 1.5 Hz, time-domain cleaning and outlier removal were performed using AMICA auto rejection~\cite{Palmer2011-zs}. Eye artifacts were removed using the ICLabel toolbox applied to the results from an AMICA~\cite{Pion-Tonachini2019-fy}. For this, ICLabel's popularity classifier was used, meaning that all components having the highest probability for the eye class were projected out of the sensor data.

For all three conditions, ERPs were extracted from band-pass filtered (0.1 Hz to 15 Hz) activity at electrode \textit{FCz}. These ERPs were obtained from -1000 ms to 500 ms around the movement onset. Trials were excluded in line with the removal for classifier training, see section~\ref{eeg_methods}. For all muscle-controlled trials without EMS, movement onset was defined as the time of the tap on the touchscreen minus the `EMG-delay' defined in section~\ref{BCI}. For all EMS-controlled trials, movement onset was defined as EMS onset. To keep only EMS-controlled trials where the stimulation resulted in an `immediate' screen tap, first all trials were rejected where there was no screen tap in the 350 ms following EMS. Next, `extreme outlier removal' was conducted on the delta between EMS and the screen tap using Tukey's method based on the inter-quartile range ~\cite{Tukey1949-sl}. In total, 32.9 (SD = 30.3) trials were rejected on average per participant.

% To analyze prediction error ERPs following the movement onset, potentially reflecting a disruptive experience impacting SoA, a linear mixed effects model was fit at each time point of the ERP using the `lme4' package~\cite{Bates2015-bh}. The model `ERP\_sample $\sim$ condition + (1|participantID)' was fit with condition reflecting trials belonging to INTENTION, INVOLUNTARY, or AUGMENTED. Prior to model fitting, all trials were aggregated per participant and condition. A test statistic was obtained by calculating likelihood-ratio tests comparing the full model as specified above against the null model `ERP\_sample $\sim$ 1 + (1|participantID)'. All parameters were estimated by maximum likelihood estimation~\cite{Pinheiro2006-bk}. P-values were corrected for multiple comparisons using \textit{false discovery rate}~\cite{Benjamini1995-cw} at alpha = .05.

% Since the AUGMENTED condition contains both, `EMS-controlled' trials, where the classifier detected an intent to interact, as well as `muscle-controlled' trials, where the classifier failed to detect the user's intent and the user carried out the tap without EMS, we split these two trial categories. We hypothesized that `EMS-controlled' trials in AUGMENTED \textit{differ} from trials in INVOLUNTARY. The latter should elicit a stronger prediction error than the former since no impact of volition is present in INVOLUNTARY, which could potentially moderate the ERP. Similarly, we hypothesized that `muscle-controlled' trials in AUGMENTED should \textit{not} differ from INTENTION trials. To test these hypotheses, permutation t-tests were conducted on aggregated data using MNE-python~\cite{Gramfort2013-fa} contrasting `EMS-controlled' trials in AUGMENTED with INVOLUNTARY as well as `muscle-controlled' trials in AUGMENTED with INTENTION. The tests were conducted for the time window from 150-250 ms post movement-- or EMS-onset event. In short, these tests were conducted to investigate the moderating role of the user acting at their own volition in AUGMENTED.

\subsection{Measures of Agency Experience: Intentional Binding, Question \& Interview}% adapt title?
% To assess if SoA was preserved, an intentional binding measure, a standardized question, and an interview following the AUGMENTED block were investigated. 
After each condition participants were prompted to rate their experienced agency on a 7-point Likert scale with the statement ``It felt like I was in control of the movements during the task.'', the item was copied from~\cite{Bergstrom2022-fb}. Following AUGMENTED we interviewed users about their experience working with the system. After prompting users to recall their experience and summarize what their task had been, we set the focus to the tapping movement and asked them to ignore the time estimation task in the questions that followed. We entered the open part of the interview by asking: ``What did the system do?'' followed up by ``What was the difference between the three conditions''. After some time, and depending on their answers, we reset the focus to the AUGMENTED condition and asked ``How often was the system active?'' followed by ``What do you think caused the actions of the system?''. This was then followed up by an `open' interview in which we frequently asked `how' and `why' questions to inquire about the user's experience.

We analyzed the interviews by loosely following~\citet{Mayring2015-pp}. All interviews were manually transcribed and translated to English using DeepL\footnote{https://www.deepl.com/translator} (DeepL SE, Cologne, Germany). Before screening the texts, two experts clustered the responses into 3 clusters: First, ``Functionality'' referred to what participants attributed the source of the stimulation. Next, we clustered responses according to ``Guessed percentage'' of correct interaction, i.e., participants' estimate of how well the system was aligned with their intention. For the cluster ``Correct Interaction'', we specifically queried participants to recall the moments where the stimulation felt in line with their intention to move, then we clustered their responses into sentiments with positive and negative valence.

\subsection{Statistical Analyses}
To confirm and demonstrate the discriminative power of the EEG features, we plotted the amplitude time course of electrode Cz between \textit{pre-move} and \textit{idle} epochs. Electrode Cz was chosen for exemplary presentation, as it is frequently reported in studies on the RP, since it captures neural activity originating from the sensorimotor cortex. Next, the slope coefficients were extracted and a paired t-test was conducted. Ultimately, to assess the classifier performance, we calculated the F1 score, i.e. the harmonic mean of precision and recall, and plotted the ROC, see figure~\ref{fig:EEG_results}a to c.

@@

In line with the literature on intentional binding, we hypothesized that the time intervals should be underestimated for the INTENTION and the AUGMENTED conditions. This should not be the case for the INVOLUNTARY condition where users had no intention to move. Hence, when binding occurs, the intervals should be underestimated, hinting at a higher SoA.

To test this, we fitted a linear mixed effects model with \textit{condition} (INTENTION, INVOLUNTARY, AUGMENTED) as a fixed effect and `participantID' as a random effect on aggregated data. The model was specified as `time interval $\sim$ condition + (1|participantID)' and fit using the `lme4' package~\cite{Bates2015-bh}. A test statistic was obtained by calculating likelihood-ratio tests comparing the full model as specified above against the null model `time interval $\sim$ 1 + (1|participantID)'. All parameters were estimated by maximum likelihood estimation~\cite{Pinheiro2006-bk}. We computed post-hoc pair-wise tests for `condition' corrected for multiple comparisons (Tukey method) using the emmeans package~\cite{Lenth2020-xk}

% We included all single trial data points in the analyses and did not aggregate before fitting the model as fitting the random effect `participantID' corrects for the correlation of the repeated measurements within participants.

Next, we hypothesized that subjective ratings of control over the tap movement are comparable between INTENTION and AUGMENTED and lower in the INVOLUNTARY condition. Again, we fitted a linear mixed effects model with \textit{condition} as a fixed effect and participant ID as a random effect. Coefficients were assessed in the same way as for the intentional binding parameter above.

In short, we tested two main hypotheses with regard to the agency experience in our three experimental conditions: (1) participants underestimate the tone delay when acting intentionally. Adding EMS in line with participants' intention to move does not affect this underestimation. (2) The subjective feeling of control is comparable between INTENTION and AUGMENTED conditions. INVOLUNTARY should decrease the feeling of control significantly. Additionally, we report the clustered interviews anecdotally.

To analyze prediction error ERPs following the movement onset, potentially reflecting a disruptive experience impacting SoA, a linear mixed effects model was fit at each time point of the ERP using the `lme4' package~\cite{Bates2015-bh}. The model `ERP\_sample $\sim$ condition + (1|participantID)' was fit with condition reflecting trials belonging to INTENTION, INVOLUNTARY, or AUGMENTED. Prior to model fitting, all trials were aggregated per participant and condition. A test statistic was obtained by calculating likelihood-ratio tests comparing the full model as specified above against the null model `ERP\_sample $\sim$ 1 + (1|participantID)'. All parameters were estimated by maximum likelihood estimation~\cite{Pinheiro2006-bk}. P-values were corrected for multiple comparisons using \textit{false discovery rate}~\cite{Benjamini1995-cw} at $\alpha = .05$.

Since the AUGMENTED condition contains both, `EMS-controlled' trials, where the classifier detected an intent to interact, as well as `muscle-controlled' trials, where the classifier failed to detect the user's intent and the user carried out the tap without EMS, we split these two trial categories. We hypothesized that `EMS-controlled' trials in AUGMENTED \textit{differ} from trials in INVOLUNTARY. The latter should elicit a stronger prediction error than the former since no impact of volition is present in INVOLUNTARY, which could potentially moderate the ERP. Similarly, we hypothesized that `muscle-controlled' trials in AUGMENTED should \textit{not} differ from INTENTION trials. To test these hypotheses, permutation t-tests were conducted on aggregated data using MNE-python~\cite{Gramfort2013-fa} contrasting `EMS-controlled' trials in AUGMENTED with INVOLUNTARY as well as `muscle-controlled' trials in AUGMENTED with INTENTION. The tests were conducted for the time window from 150-250 ms post movement-- or EMS-onset event. In short, these tests were conducted to investigate the moderating role of the user acting at their own volition in AUGMENTED.

### 2024-10-04 submission_JNE/source/input/6_results.tex
page and jne submission

\section{Results}
We first report the performance of the classification system, as the system performance greatly impacts the subjective experience of the motor augmentation.

\subsection{Behavior: Intentional Binding}
In the intentional binding task, i.e. the estimation of the time interval between tap and tone, participants generally underestimated the average real delay (350ms) in all three conditions (INTENTION M = 187.5, SD = 139; EXTERNAL M = 208.6, SD = 140.1, AUGMENTED M = 196.0, SD = 138.2). The condition influenced the estimation (${\chi^{2}_{(2)}} = 17.1, p < .001$), see figure~\ref{fig:results}a. Post-hoc tests revealed that the underestimation was less pronounced in EXTERNAL as compared to INTENTION ($beta = 22.6, p < 0.001$). This was also the case in EXTERNAL, with less pronounced underestimations of the time interval between tap and tone compared to AUGMENTED ($beta = 15.3, p = 0.018$). No difference was observed between INTENTION and AUGMENTED.

\subsection{Subjective Ratings \& Reports}
The subjective rating of control differed between conditions (${\chi^{2}_{(2)}} = 43.7, p < .001$), see figure~\ref{fig:results}b. Post-hoc tests revealed that participants rated their motion control higher in INTENTION compared to EXTERNAL ($beta = 4.8, p < .0001$), and higher in INTENTION compared to AUGMENTED ($beta = 3.1, p < .0001$). Further, higher control was observed in AUGMENTED compared to EXTERNAL ($beta = -2.9, p = .006$).

\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth]{figures/subjective_results.pdf}
    \caption{(a) Temporal binding effect over conditions. With an average presented time interval of 350ms, lower values in the estimated time interval indicate an underestimation, i.e., temporal binding. (b) Subjective ratings of control across conditions. Significance labels obtained from post-hoc tests on estimated marginal means.}
    \label{fig:results}
\end{figure}

@@

Some perceived the stimulation as externally triggered, yet partially responsive to their choices, resulting in a sentiment captured by, ``I think that the stimulation in the third block was partly involuntary, but partly as if it was following my decision/choice.''. One participant noted that they felt that ``somehow the information is in my arm''. Another one briefly considered the involvement of their brain waves in controlling the stimulation, but expressed doubts about this possibility, stating, ``But I don't think that is the case.''.

@@

>>>>

\section{Results}
% We first report the performance of the classification system, as the system performance greatly impacts the subjective experience of the motor augmentation.

% \subsection{Behavior: Intentional Binding}
In the intentional binding task, i.e. the estimation of the time interval between tap and tone, participants generally underestimated the average real delay (350ms) in all three conditions (INTENTION M = -160.7 ms, SD = 88.1; INVOLUNTARY M = -135.3, SD = 110.3, AUGMENTED M = -162.4, SD = 113.8). The underestimation was not affected by the condition, see~\ref{fig:ib_task}.

% The condition influenced the estimation (${\chi^{2}_{(2)}} = 17.1, p < .001$), see figure~\ref{fig:results}a. Post-hoc tests revealed that the underestimation was less pronounced in INVOLUNTARY as compared to INTENTION ($beta = 22.6, p < 0.001$). This was also the case in INVOLUNTARY, with less pronounced underestimations of the time interval between tap and tone compared to AUGMENTED ($beta = 15.3, p = 0.018$). No difference was observed between INTENTION and AUGMENTED.

\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth]{figures/ib_task.pdf}
    \caption{Difference in time estimation from real delay in intentional binding task for the three experimental conditions (left). Negative values indicate a temporal compression, i.e., temporal binding. Right side: Trials in the AUGMENTED condition split in EMS-- and self-executed trials.}
    \label{fig:ib_task}
\end{figure}

\subsection{Subjective Ratings \& Reports}
The subjective rating of control differed between conditions (${\chi^{2}_{(2)}} = 43.7, p < .001$), see figure~\ref{fig:loc}b. Post-hoc tests revealed that participants rated their level of control higher in INTENTION compared to INVOLUNTARY ($beta = 4.8, p < .0001$), and higher in INTENTION compared to AUGMENTED ($beta = 3.1, p < .0001$). Further, higher control was observed in AUGMENTED compared to INVOLUNTARY ($beta = -2.9, p = .006$).

% \begin{figure}[h]
%     \centering
%     \includegraphics[width=\columnwidth]{figures/subjective_results.pdf}
%     \caption{(a) Temporal binding effect over conditions. With an average presented time interval of 350ms, lower values in the estimated time interval indicate an underestimation, i.e., temporal binding. (b) Subjective ratings of control across conditions. Significance labels obtained from post-hoc tests on estimated marginal means.}
%     \label{fig:results}
% \end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=.5\columnwidth]{figures/level_of_control.pdf}
    \caption{Subjective ratings of control across conditions. Significance labels obtained from post-hoc tests on estimated marginal means.}
    \label{fig:loc}
\end{figure}

@@

Some perceived the stimulation as externally triggered, yet partially responsive to their choices, resulting in a sentiment captured by, ``I think that the stimulation in the third block was partly random, but partly as if it was following my decision/choice.''. One participant noted that they felt that ``somehow the information is in my arm''. Another one briefly considered the involvement of their brain waves in controlling the stimulation, but expressed doubts about this possibility, stating, ``But I don't think that is the case.''.

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/eeg_results_new.pdf}
    \caption{(a) Grand average event-related potential (ERP) at electrode Cz of pre-movement EEG data epochs preceding the last second before a movement, idle epochs of the same trial are plotted alongside but originate from a different time window, see section~\ref{eeg_methods}; (b) Bottom: Slope features for both idle and pre-movement classes at electrode Cz. Top: Scalp maps of slope values for both classes and all channels. (c): Mean and per participant ROC curves. Dotted line at 15 \% false positive rate indicate the selected threshold for the real-time application.}
    \label{fig:EEG_results}
\end{figure*}

\subsubsection{Post-hoc analyses of classifier performance}

In line with the readiness potential a negative going deflection over the last second preceding the movement onset was present in both INTENTION and AUGMENTED, but not in INVOLUNTARY, see figure~\ref{fig:erp}a. Furthermore, ERPs differed between the three conditions in the time window from 210 ms -- 250 ms following movement onset ($\chi^{2}_{(2} = 10.6, p = .03$ at 220 ms). INVOLUNTARY exhibited a strong negativity peak (strongest among the three conditions), peaking at around -10 $\mu$V at 210 ms after movement onset. A weaker peak was observed for AUGMENTED (-6 $\mu$V at 210 ms) and even more so for INTENTION (-2 $\mu$V at 150 ms).

\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth]{figures/PE_erp.pdf}
    \caption{ERPs at electrode FCz from -1000 before, to 600 ms after the tapping movement onset. Top (a): INTENTION, INVOLUNTARY and AUGMENTED trials. Red bar on top indicate main effect at $p<.05$. Bottom (b): Difference ERPs at FCz of the `muscle-controlled' trials in AUGMENTED minus INTENTION as well as EMS trials in AUGMENTED minus INVOLUNTARY.}
    \label{fig:erp}
\end{figure}

When correcting for the influence of the EMS by subtracting INVOLUNTARY from EMS trials in AUGMENTED an INTENTION from `muscle-controlled' trials in AUGMENTED we observed no differences. Visually, it appears that trials in both INVOLUNTARY and INTENTION condition exhibit a stronger negativity in the post movement time window of interest than respective EMS and `muscle-controlled' trials in AUGMENTED, see the positive deflections peaking at 300 ms post movement onset in figure~\ref{fig:erp}b.
